----
title: "Reproducible Research Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## **Loading all the packages used**

```{r loading,warning=FALSE,message=FALSE}
library(dplyr)
library(knitr)
library(ggplot2)
library(Hmisc)
```


# **Loading and preprocessing the data**

## **Loading**
### Reading the file **activity.csv** into *activity_data*
```{r}
activity_data <- read.csv('./Fitbit/activity.csv')
```

## **Processing**
### Removing NAs using *na.omit()* and storing in *activity_data_na_omitted*
```{r}
activity_data_na_omitted <- na.omit(activity_data)
```


# **Q. What is mean total number of steps taken per day?**

## **Total number of steps taken per day**
```{r}
daily_steps <- group_by(activity_data_na_omitted,date) %>% summarise(Total.Steps = sum(steps))
head(daily_steps)
```

## **Histogram of the total number of steps taken each day**
```{r}
hist(daily_steps$Total.Steps,main="Total number of steps per day",xlab="Total number of steps in a day")
```

## **Mean and Median of total steps per day**

### Mean
```{r}
mean(daily_steps$Total.Steps)
```

### Median
```{r}
median(daily_steps$Total.Steps)
```

# **Q. What is the average daily activity pattern?**

## **Average number of steps taken vs Interval**
```{r}
step_interval <- aggregate(steps ~ interval, activity_data_na_omitted, mean)
plot(step_interval$interval, step_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")
```

##  **Interval with maximum number of steps**
```{r}
step_interval[which.max(step_interval$steps),]
```


# **Imputing missing values**

##  **Total number of missing values in the dataset**
```{r}
sum(is.na(activity_data))
```

## **Imputing steps using impute with mean**
```{r}
activity_data$Imputed.Steps <- with(activity_data,impute(steps,mean))
```

## **Creating dataset *imputed_daily_steps* **
```{r}
imputed_steps <- transform(activity_data,steps=Imputed.Steps)
``` 

## **Histogram of the total number of steps taken each day**
```{r}
imputed_daily_steps <- aggregate(steps ~ date, imputed_steps, sum)
hist(imputed_daily_steps$steps, main="Total number of steps per day (imputed)", 
     xlab="Total number of steps in a day")
```

## **Mean and Median of total steps per day compared**

### Mean *Imputed*
```{r}
mean(imputed_daily_steps$steps)
```

### Mean
```{r}
mean(daily_steps$Total.Steps)
```

### Median *Imputed*
```{r}
median(imputed_daily_steps$steps)
```
### Median
```{r}
median(daily_steps$Total.Steps)
```

# **Q. Are there differences in activity patterns between weekdays and weekends?**

## **Creating a new factor variable in the dataset with two levels – “weekday” and “weekend”**
```{r}
imputed_steps['Day.Type'] <- weekdays(as.Date(imputed_steps$date))
imputed_steps$Day.Type[imputed_steps$Day.Type %in% c('Saturday','Sunday') ] <- "Weekend"
imputed_steps$Day.Type[imputed_steps$Day.Type != "Weekend"] <- "Weekday"
imputed_steps$Day.Type <- as.factor(imputed_steps$Day.Type)
head(imputed_steps)
```

## **Plotting Steps vs Interval for Weekends and Weekdays**

```{r}
imputed_steps_interval <- aggregate(steps ~ interval + Day.Type, imputed_steps, mean)
qplot(interval,steps,data=imputed_steps_interval,type = 'l',geom=c("line"),xlab = "Interval",ylab = "Number of steps",main = "") +facet_wrap(~ Day.Type, ncol = 1)
```
